---
title: "Homework 1"
author: "Dylan Zumar"
date: "02/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

## Task 1

```{r mh_distance_function}
mh_distance <- function(x, y) {
  #Check logical, numeric, character
  if(!is.logical(x) & !is.numeric(x) & !is.character(x)) {
    warning("x is not logical, numeric, or character")
    return(-1)
  }
  if(!is.logical(y) & !is.numeric(y) & !is.character(y)) {
    warning("y is not logical, numeric, or character")
    return(-1)
  }
  #Check lengths
  if (is.numeric(x) & is.numeric(y)) {
    if (nchar(abs(x)) != nchar(abs(y))) {
      warning("x and y have different lengths") {
        return(-1)
      }
    }
  }
  if (is.character(x) & is.character(y)) {
    if (nchar(x) != nchar(y)) {
      warning("x and y have different lengths")
      return(-1)
    }
  }
  #Check decimals
  if (is.numeric(x) & (round(x) != x)) {
    warning("x is a decimal")
    return(-1)
  }
  if (is.numeric(y) & (round(y) != y)) {
    warning("y is a decimal")
    return(-1)
  }
  #Check NA, NaN, Inf, -Inf
  if (is.na(x) | is.na(y) | is.nan(x) | is.nan(y) | is.infinite(x)
      | is.infinite(y)) {
    warning("x or y is NA, NaN, Inf, or -Inf")
    return(-1)
  }
  
}
```


## Task 2

<i>
Be sure to remove chunk option `eval=FALSE` in the below code chunks
once you create `mh_distance()`. Do not include this italicized
text in your final submission.
</i>

**Initial test cases** (return a non-negative modified Hamming distance)

```{r eval=FALSE}
mh_distance(x = "abc", y = "abc")
mh_distance(x = T, y = FALSE)
mh_distance(x = "523890", y = "752839")
mh_distance(x = 2341, y = 2350)
```

Added test cases that return a non-negative modified Hamming distance result.

```{r}
# additional test cases go here
```

**Initial test cases** (return values of `-1`)

```{r eval=FALSE}
mh_distance(x = 52, y = 113)
mh_distance(x = "swimming", y = "winning")
mh_distance(x = NA, y = TRUE)
mh_distance(x = 1.5, y = 2.5)
```

Added test cases that return a value of `-1`.

```{r}
# additional test cases go here
```


## Task 3

Consider the pair of vectors `s` and `w` given below.

```{r}
s <- c(26, 50123, 456.12, 8, 0)
w <- c(22, 50000, 451.00, 88, 0)
```

```{r}
# loop goes here
```


## Task 4

Brief write-up goes here...
